### 2014-06-01 text/askalon.tex
separate desc

\paragraph{ASKALON/AGWL.} The ASKALON Problem Solving Environment (PSE)
\cite{fahringer2014} is developed at the University of Innsbruck.
ASKALON is described as a Scientific PSE capable of composing, hosting,
executing, sharing and reusing scientific workflows written in the Abstract
Grid Workflow Language (AGWL). AGWL is an XML-based workflow language that is
capable of expressing scientific application components chains at a high-level
of abstraction. A concrete workflow consists of the abstract workflow combined
with the resources and target execution environment binding the workflow. The
graphical interface of AGWL is based upon the UML (Unified Modeling Language)
diagrams.\km{Todo: current developments and advances.}