### 2014-06-01 text/galaxy.tex

\paragraph{Galaxy Portal/JSON} Galaxy~\cite{blankenberg-vonkuster-etal:2010} is a
SWE developed at the Penn State University, Pennsilvania. It provides a
webportal based environment to build complex pipelines from an available
repository of tools (\textit{http://g2.bx.psu.edu}).  Galaxy is currently used
mainly by bioinformatics community. Users can extend the tool repository
provided by Galaxy using their own Galaxy server. One of the major use of
Galaxy is to achieve a mix and match of data analysis tools in order to analyze
a vast repository of genome based database. Galaxy offers a web-based,
interactive platform for data analysis via reusable and reproducible workflows.
While excelling in the usability aspects, the Galaxy environment has a limited
parallelism capabilities and a small number of existing implementations for
interfacing with diverse, scalable computing infrastructures. The user can draw
workflows in which multiple copies of a task are created to operate on multiple
inputs at once, but this is not a scalable approach.