### 2014-06-01 text/makeflow.tex

\paragraph{Makeflow/make-like spec} Makeflow~\cite{makeflow} is a make-like workflow platform.
Makeflow addresses the often-encountered problem of software and library
dependencies on workflow tasks execution for heterogeneous environments.
Makeflow offers techniques to ``link" and package the application execution
elements such as executables, libraries and configuration parameters. This
improves the durability of a workflow and decreases the efforts required to
maintain a working version of an application encoded as a workflow.

%Internal representation of a Makeflow workflow is in...
%Its expressivity features include...
%Makeflow's runtime features include...
%