### 2014-06-01 text/moteur.tex

\paragraph{MOTEUR/scufl, GWENDIA} MOTEUR \cite{glatard-montagnat-etal:2008} is a SWE
developed at the I3S laboratory, University of Nice at Sophia Antipolis (UNS).
MOTEUR facilitates executing workflows consisting of remote services. It
exploits service parallelism at workflow level and data parallelism for
multiple datasets.

MOTEUR has sophisticated language-level expressions for data management
mechanisms within a workflow definition. It provides a diverse set of
parallel iterators that are well adapted for different combinations of input datasets
at the processor. An iteration strategy allows different combinations of the
data arriving at a processor along its multiple inputs. Running legacy
application code on distributed systems has been eased by MOTEUR by providing a
generic grid-service wrapper for these applications
\cite{glatard-montagnat-etal:2008a}. Services thus created can be readily
deployed by the MOTEUR runtime. MOTEUR workflows are written in the scufl
language and are compatible with those written for the Taverna SWE. It also
supports an extension of scufl language called the Gwendia~\cite{gwendia}
language, developed at the same laboratory. 
%