### 2014-06-01 text/pgrade.tex

\paragraph{The P-GRADE Portal/--} The P-GRADE Portal~\cite{pgrade2011}
developed at the MTA SZTAKI Computation and Automation Institute, Budapest,
provides an access to grid computational and storage resources from a web
browser through a specialized workflow oriented grid portal. A workflow
development environment interfaces the users through this web based portal. The
portal interfaces the multiple grid computing middleware and storage resources
as a backend component of the environment. A highlight of P-GRADE portal is
that it provides access to multiple grids and scientific communities in a
single environment leading to a highly collaborative inter-community,
inter-grid research environment. An older version of P-GRADE used the
condor-Dagman as workflow engine while the recent version uses its own in-house
developed workflow engine called Xen. \km{Investigate cloud and GPU capabilities}