### 2014-06-01 text/triana.tex

\paragraph{Triana/XML} Triana \cite{taylor-shields-etal:2003} is a java-based
SWE developed at the Cardiff University. It consists of a workflow composition
GUI, a framework to integrate workflow with the different execution
environments and middlewares and a set of predefined services written as java
components. A Triana workflow is expressed as an XML grammar similar to a WSDL
description for webservices \cite{taylor-wand-etal:2005}. A highlight of
Triana is that it provides ready-to-use interface for a wide variety of
middleware and distributed infrastructures including P2P, grids and
web-services frameworks. A Triana workflow provides dataflow as well as control
flow support. Triana allows conditional branching and looping in its workflows.