### 2014-06-01 text/wings.tex

\paragraph{Wings/--} The Workflow instance generation and specialization
(Wings) provides automatic workflow validation and generation based on
high-level workflow templates~\cite{wings2013}. The system receives high-level
workflow specification in the form of Pegasus/DAX workflows and validates them
for correct coordination based on data and resource constraints. It generates
workflows by selecting components, data sets, and parameters in three stages
while eliminating candidates that are not viable. This approach is useful in
the context of desiging complex computational experiments by domain scientists
in that it eases the process of workflow level debugging. Wings uses W3C's OWL,
Resource Description Framework (RDF), and Semantic Web Rule Language (SWRL) to
represent workflows and their associated constraints.

% How does Wings achieve the translation between abstract and concrete workflows?
% What are the constraints under which Wings works?
% How is Wings implemented?

%