### 2014-06-02 paper.tex
refactor

% v2-acmsmall-sample.tex, dated March 6 2012
% This is a sample file for ACM small trim journals
%
% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
% (c) 2010 Association for Computing Machinery (ACM)
%
% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
% Users can also go through the FAQs available on the journal's submission webpage.
%
% Steps to compile: latex, bibtex, latex latex
%
% For tracking purposes => this is v1.3 - March 2012

\documentclass[prodmode,acmtecs]{acmsmall} % Aptara syntax

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\usepackage{cite}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{mindmap,trees}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

\usepackage{setspace}
%\doublespacing
\linespread{2}

% Metadata Information
\acmVolume{9}
\acmNumber{4}
\acmArticle{39}
\acmYear{2010}
\acmMonth{3}

\newcommand{\jm}[1]{
    \noindent{\textcolor{red}{\bf\em #1 -- Johan}}
}
\newcommand{\km}[1]{
    \noindent{\textcolor{blue}{\bf\em #1 -- Ketan}}
}

% Document starts
\begin{document}

% Page heads
\markboth{K. Maheshwari et al.}{A Survey of Scientific Workflow Management Systems}

% Title portion
\title{State of the Art in Scientific Workflow Systems}
\author{KETAN MAHESHWARI
\affil{University of Nice, Sophia Antipolis}
JOHAN MONTAGNAT
\affil{CNRS}
}
% NOTE! Affiliations placed here should be for the institution where the
%       BULK of the research was done. If the author has gone to a new
%       institution, before publication, the (above) affiliation should NOT be changed.
%       The authors 'current' address may be given in the "Author's addresses:" block (below).
%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
%       currently affiliated with NASA.

\begin{abstract}

\km{TODO: Expand in the end}

In this work we conduct a literature survey of the key features of the
contemporary scientific workflow environments. We define the criteria for the
survey based chiefly on user needs arising from diverse applications and
infrastructure. An overview of studied scientific workflow environments is
presented followed by the identification and description of the features offered by
each of the environment. A tabulated summary for each feature class gives a
high-level view of the offered features. Finally, a taxonomy is proposed
placing these features into respective places in a hierarchy.

\jm{Test comment}

\end{abstract}

%\category{C.2.2}{Computer-Communication Networks}{Network Protocols}

%\terms{Design, Algorithms, Performance}

%\keywords{Wireless sensor networks, media access control, multi-channel, radio interference, time synchronization}

\acmformat{Ketan Maheshwari, Johan Montagnat
, 2014. A survey of scientific workflow management systems.}
% At a minimum you need to supply the author names, year and a title.
% IMPORTANT:
% Full first names whenever they are known, surname last, followed by a period.
% In the case of two authors, 'and' is placed between them.
% In the case of three or more authors, the serial comma is used, that is, all author names
% except the last one but including the penultimate author's name are followed by a comma,
% and then 'and' is placed before the final author's name.
% If only first and middle initials are known, then each initial
% is followed by a period and they are separated by a space.
% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.

\begin{bottomstuff}
This work is supported by the ANR-CNRS, under
grant contract CNRS-0435060.

Author's addresses: Ketan Maheshwari, Argonne National Laboratory; Johan Montagnat, I3S, Sophia Antipolis
\end{bottomstuff}

\maketitle

%Workflows considered:
% MOTEUR2
% Taverna
% Swift
% Pegasus
% Triana
% Galaxy
% Makeflow
% Kepler/Ptolemy
% Dryad
% Science pipes
% Spark
% Linda
% DaGue
% Wings
% PGRADE Portal

\section{Introduction}
\label{sec:intro}
\vspace*{0.5cm}\hspace{1cm} Scientific Workflow Environments (SWE)
are, in principle similar to the traditional programming environments
(programming language, interpreter or compiler, and runtime) albeit with a
larger scale of data and a higher abstraction of computation. The similarity is
such that scripting languages are sometimes advocated for programming at this
level. However, traditional scripts do not deal with parallelism and dataflow
aspects of workflow paradigm which makes them a strong contenders for the
purpose of high-level application programming. Computation in science is
dominated by computation for data analysis via one or more mathematical or
scientific model execution. These models of computation are composed of
multiple stages where each stage executes a program requiring input parameters
and data while producing results in the form of data which in turn acts as
input to the next stage. SWEs can ease these activities by supporting the
expression, execution and automation.
% Goals and benefits of a workflow management system
The goals and benefits of an SWE are as follows:

\begin{enumerate}
\item \textbf{Bulk Work.}
Workflows are important because they enable to perform computational work in
bulk. This could be bulk data processing or running multiple distinct
processing tasks on a single set of data. In addition to support such
work, workflow environments provide oppurtunities for giving structure and
order to a set of tasks and often with sufficient expressivity, choices and
selections to carry out a bulk of work selectively. Some of the examples of
this approach are \emph{parameter sweeps} where tasks are run over a series of
parameter values, \emph{ensemble studies} where tasks are carried out over
discrete groups of parameters over a space.

\item \textbf{User Interaction.} Adaptation to new methods of computing is not
easy for domain science users. Such users often need special assistance in
order to translate their computation requirements into concrete computing
entities. SWEs are an ideal platform enabling users to perform many tasks
conveniently. These tasks include visually formulating an analysis, composing
and marshalling of processes, browsing and selecting resulting data,
visualizing and sharing final results.

\item \textbf{Resource Access and Utilization.} Computational platforms and
access modalities change rapidly with upgradation of existing technologies and
adaptation of new techniques. Clouds are an example of new techniques
of computational platform access while GPGPUs are example of upgradation of
existing processing environments. Additionally, with the advent of fast and
dedicated networks such as ESNet~\cite{esnet} and federated remote
visualization resources are gaining prominence among scientific users. SWEs
play important role as platforms to give transparent and hassle free access to
such novel and multiple environments to users by hiding the technicalities of
access methods and usage and configuration related nuances.

\item \textbf{Repeatable Execution.} Scientific users often require a
repeatable execution of same application(with same parameters and input data)
for reproduction of results and demonstration of methods. SWEs can record and
save workflows, associated parameters, input data and historical results to
aid in such activities. Additionally, SWEs facilitate an incremental
development of these methods leading to augmentation of past efforts to current
and future developments.

\item \textbf{Provenance and MetaData.} Historical data and metadata has been
crucial entities in many research fields. They help connecting results from
a multitude of research points of view. A provenance trace of a computation can
help checkpoint methods and reproduce and modify computation fully or
partially. Automated observations of deviations in scientific phenomena from
provenance data is one of the key triggers for new scientific discoveries.
Fields of studies involving genomic and protiens are examples of domains where
provenance and metadata plays key roles. SWEs can be integrated or extended
with metadata and provenance techniques rendering powerful computing and
traceability features. In fact, most SWEs of today have separate modules or
extensions for provenance and metadata management.

\item \textbf{Sharing and Collaboration.} SWEs are often targetted
towards a particular community of users. Dissemination of methods and results
to communities and collaborators are some of the activities that SWEs make
convenient for end-users. For instance, the
Galaxy~\cite{blankenberg-vonkuster-etal:2010} workflow system is targetted
specifically for the bioinformatics users encouraging them to collaborate over
problems and share results. Another such example is Taverna~\cite{taverna2013}
which is geared towards sharing and collaborative workflow oriented scientific
computation enabling users to discover new services introduced to a community.
Sharing and collaboration of applications and related data have become
ubiquitious with the rise of social web. Initiatives such as
\emph{myExperiment}~\cite{deroure-goble-etal:2008} and recently
\emph{gitHub}~\cite{github-web} have enabled easy sharing and collaboration of
reproducible workflows on the social web.

\end{enumerate}

The above described goals and benefits also form a basis for the need for a
SWE. The need for the use of SWE can arise as a result of one or more of such
requirements. Traditional programming platforms are not suitable to address
these requirements as they operate at a different level of abstraction. SWEs
differ in many respects among themselves as well as with respect to traditional
programming environments. Consequently, a choice of the suitable SWE must be
made by a researcher. The following list identifies the major differences
between traditional and the SWE programming paradigm:

\begin{enumerate}
%
\item \textbf{Scalability and Optimization.} SWEs are more likely to involve
extensive usage of distributed resources than conventional programs. Owing
to the large-scale experiments, for which scientific users turn to automation,
their demands for computational power often exceeds that of a single personal
computer. These properties make them ideal candidates for a large-scale
distributed computing environment. The case for suitability of SWEs on
large-scale distributed infrastructures has been well-argued upon
in~\cite{jha-katz-etal:2009}. In such a scenario, taking explicit steps to
ensure scalable and optimized performance is vital. Additionally, most SWEs are
capable of executing workflows at small scale with minimal tuning. This
multi-scalability enables quick and easy conceptualization and prototyping of
new workflows.
%
\item \textbf{Data Management.} The scale of data transfer is much higher in
SWEs from one execution site to others as well as among storage resources while
it is lower (ranging from variables to small data structures) in programs.
Owing to such scales, applications often need access to special data movement
and access and movement protocols such as GridFTP, Amazon S3, NAS, etc. SWEs are often
shipped with capabilities to handle such protocols.
%

\item \textbf{Interface to Distributed Computing Infrastructures (DCIs).}
Supporting an interface to a distributed system is a complex task from the
point of view of an SWE. The distributed nature of resource utilization and
asynchronous execution of workflow requires a set of interdependent parameters
to be configured correctly. Modern computing infrastructures such as Clouds,
GPUs and supercomputers with hybrid nodes (CPU + GPGPU) pose new challenges of
interface development.  As SWEs involve running many tasks over an
infrastructure it becomes crucial to interface with these systems such that the
tasks are scheduled in an optimal manner. Many SWEs implement
\emph{pilot-jobs}~\cite{montagnat-glatard-etal:2010,ucc2011} in order to
reserve resources via pseudo-tasks that act as placeholders for real tasks.
Additionally, decision to manage working data storage on disc, memory and cache
are some of the distinctions between a traditional programming and workflow
platforms in the context of distributed and scalable computation.

\item \textbf{Expressiveness.} Programming languages are inherently control
flow oriented where data flow is a secondary aspect. In fact, the
programmer must explicitly make sure that correct inputs are available
to instructions and are in the right format. A program proceeds as the
execution of its statements from beginning to end and completes its
execution in a predefined sequence. However, in SWEs, the flow is often
determined by the data availability for execution. This model of computation is
known as dataflow computing. A workflow thus does not follow a fixed order of
execution. Different execution sequences are possible which are dominated by
dataflow. The unit of work in SWEs is at a higher level of granularity. A unit
of execution is coarse-grained, and often termed as ``processor'', ``task'',
``service'' or ``actor'' in the SWE community. This notion of high-level
granularity drives the depth to which a workflow can be expressed in a given
language form. In the rest of the paper, we refer to this unit of execution of a
workflow system as \emph{processor}. On the other hand, a unit of execution in
most programming languages is a much simpler and finer executable statement
that accomplishes a small task. This presents a whole new set of
requirements for expressing the processors and how they should interact with
computational resources and operate upon data.

%
\item \textbf{Usability.} SWEs predominantly serve as a tool for the domain
experts to utilize the computational power. This makes it essential for
them to have a friendly, and intuitive user-interface for novice and
non-expert users. For this reason, SWE user interfaces are dominated by the
visual paradigm of programming. On the other hand, as the sophistication of
expression and user expertise increases, the visual environment temds to get
limited in capabilities it can offer. Thus, most SWEs have a text-based
composition environment albeit with visually assisting programming environments
such as IDEs. However, these textual environments are seldom to the extent that
a non-expert user can graphically compose programs. While a graphical interface
enhances usability, a text-based workflow environment can arguably give better
control to an experienced user. Consequently, many workflow systems offer a
dual interface of graphical and text medium to compose workflows.

\end{enumerate}

We concede that the above is not an exhaustive list of differences in features.
However, it is reasonably inclusive and we believe it covers the breadth of the
aspects we discuss in the current work. We do not include other interesting SWE
features such as application predictability in this analysis as they are out of
scope of the objectives of the current paper.

\input{text/concepts}
\input{text/related}
\input{text/objectives}

\section{Overview of the Workflow Environments} \label{sec:overview}
In this section we present an overview of the SWEs studied. The paragraph
headers are organized by the name of the SWE followed by the name of internal
representation language use and support separated by a forward slash (/).  The
cases where an intermediate language is not specifically defined or conversely
where there are multiple systems supporting a generic representation are
denoted by an asterisk (*) or a hyphen (-) at the respective places.

\input{text/taverna}
\input{text/moteur}
\input{text/makeflow}
\input{text/triana}
\input{text/kepler}
\input{text/peg}
\input{text/wings}
\input{text/swift}
\input{text/askalon}
\input{text/pgrade}
\input{text/galaxy}
\input{text/others}
% Put this in analysis section. 
%The \textit{gscript}~\cite{maheshwari-montagnat:2010} workflow language
%is similar in spirit to the SwiftScript. Similar to SwiftScript, gscript adopts futures to express
%asynchronous execution in a script-based language. A Swift script 
%contains statically typed datasets that can be declared
%in the script. Unlike Swift script, gscript does provide user-defined
%strategies in order to effect data combinations in the workflow. The underlying datasets organization in Swift is
%represented using XML-representation. This means that effectively, a
%workflow is a DAG. A workflow is
%simpler and compact in gscript as its data description is embedded within a
%workflow representation. Conditional and Loop execution is supported
%using the if and foreach statements in both languages.
%
\input{text/scale}
\input{text/data}
\input{text/interface}
\input{text/lang}
\input{text/usability}

\section{Summary and Conclusions} \label{sec:sum}
% \begin{sidewaysfigure}
% \begin{center}
% \includegraphics[width=\linewidth]{2/figures/pdf/Taxonomy}
% \caption{A Taxonomy of the Features of Scientific Workflow Environments}
% \label{fig:taxonomy}
% \end{center}
% \end{sidewaysfigure}

%\begin{sidewaysfigure}
\begin{figure}
\begin{center}
\begin{tikzpicture}[]
 \path[mindmap,concept color=darkgray,text=white,font=\sf]
 node[concept] {Scientific Workflows}
 [clockwise from=0]
 child[concept color=darkgray,font=\sf] {
 node[concept] {scalability}
 [clockwise from=90]
 child [concept color=gray,font=\sf] {
 node[concept] {\small{parallelism}}
 [clockwise from=120]
 child[concept color=lightgray,text=darkgray,font=\sf] {node[concept]{data}}
 child[concept color=lightgray,text=darkgray,font=\sf] {node[concept]{\scriptsize{pipeline}}}
	 child[concept color=lightgray,text=darkgray,font=\sf] {node[concept]{\scriptsize{task}}}
% 	 child[concept color=lightgray,text=darkgray,font=\sf] {node[concept]{\small{activity}}}
 }
 child[concept color=gray,font=\sf] { node[concept] {\scriptsize{optimization}} }
 }
 child[concept color=darkgray,font=\sf] { node[concept] {\small{data management}}}
 child[concept color=darkgray,font=\sf] {
 node[concept] {\small{expressiveness}}
 [clockwise from=-45]
 child[concept color=gray,font=\sf] { node[concept] {iterators} }
 child[concept color=gray,font=\sf] { node[concept] {\small{array programming}} }
 child[concept color=gray,font=\sf] { node[concept] {\small{control structures}} }
 }
 child[concept color=darkgray,font=\sf] {
 node[concept] {usability}
 [clockwise from=-120]
 child[concept color=gray,font=\sf] { node[concept] {\scriptsize{composition}} }
 child[concept color=gray,font=\sf] { node[concept] {extendible} }
 child[concept color=gray,font=\sf] { node[concept] {reusable} }
 }
 child[concept color=darkgray,font=\sf] { node[concept] {interface to DCIs} };
\end{tikzpicture}
\caption{A Taxonomy of the Features of Scientific Workflow Environments}
\label{fig:taxonomy}
\end{center}
\end{figure}

Figure \ref{fig:taxonomy} shows a taxonomy of SWE features studied in this
paper. We studied twenty SWEs from different scientific fields developed
worldwide. Some trends as observed in the approaches of the SWEs development
are noted below:

\begin{itemize}
 \item A recent surge in distributed computing approach of Cloud Computing has
     given rise to adoption of new enactment technologies.

 \item SWEs tend to offer complementary properties which has led to several
     collaborations among the developing groups. Notable among these are the
     recent ones between Taverna and Galaxy \cite{mohamed-alaa-etal:2010} and
     Kepler and Pegasus \cite{mandal-deelman-etal:2007}.

 \item Collection oriented SWEs wherein arrays of data are treated as
     first-class entities has been seen a significant trend among the
     contemporary SWEs.

 \item A trend of increasing number of SWEs providing a script-based workflow
     composition interface solely or in combination with a visual interface in
     the form of GUI has been observed.

 \item Workflow systems originally developed for business domains have been
     introduced to scientific domains, such as BPEL and Trident.

\end{itemize}

In the present work we explore different dimensions of SWE paradigm. We study a
number of SWE implementations. While we see a number of implementations
dominant in areas suitable to their targeted applications, there is no single
SWE that could sufficiently fulfill the objectives of this work. There is a
lack of an SWE addressing the workflow expressibility and parallelism
requirements while at the same time targeting enactment within the Grid
Environment. Key features such as effective application data combinations or
array-based execution coupled with Grid centric execution are either missing or
partially available in the studied workflows.

However, so far, from our study, we find Taverna as being the closest SWE that
can be a suitable candidate for our enactments. With this motivation, we get
into studying the details of the workings of the Taverna Workflow Environment.

While many contemporary SWEs are studied in this survey, we are aware that this is not an exhaustive list. Some other SWEs have been missed or omitted, notable among them being: ICENI \cite{furmento-mayer-etal:2002}, Calcium \cite{caromel-henrio-etal:2008} and YML \cite{delannoy-emad-etal:2006}.

% Bibliography
\bibliographystyle{ACM-Reference-Format-Journals}
\bibliography{sanebib}
%\nocite{*}
\end{document}