### 2014-06-02 text/interface.tex

% == INTERFACE TO COMPUTING AND DATA INFRASTRUCTURES ==
\section{Interface to Distributed Computing Infrastructures (DCIs)} \label{sec:inter}
Workflow environments are external systems to large scale computing
infrastructures. They require an interface to connect to and use the
infrastructures. Provided the complex and distributed nature of grids, it is
not easy and practical to interact with such an infrastructure directly. Most
distributed systems have a software service layer through which it interacts
with the outside world. These interactions include accepting queries for
information and accepting requests and data for job executions. Such a layer is
typically called the \textit{middleware}. A workflow engine intending to
interact with a grid environment must provide interface to link to such
middleware layer.

Most grid computing systems today are batch oriented execution systems (with a
notable exception of GridRPC \cite{seymour-nakada-etal:2002} based systems). A
batch of ``jobs" are submitted to the system which in turn passes through some
execution queues before actually being scheduled for execution. Such systems
are prone to high execution latencies and unreliability issues. A workflow
manager must take these issues into account while interacting with such
systems.

At the resource management level, most grid computing environments today are
utilizing one or more components of the Globus Toolkit
\cite{foster-kesselman:1997} or the Condor distributed computing software
system~\cite{dagman}.

The Globus Toolkit is a software stack that provides useful functionalities for
grid computing operations. These functionalities are provided in the form of a
software stack that ranges from low-level data management tasks to the higher
level functionalities like authentication and metadata services. Any external
entity accessing the grid computing system must possess proper authentication
and sufficient authority to operate upon the infrastructure. The most popular
secure access system in place for today's grids is the public key
infrastructure (PKI) system through which a secure access to the grid is
guaranteed for a finite time period. A workflow manager must be capable of
handling the security protocols with the underlying execution infrastructure.
The duration of access to the system is also of critical importance as a
typical scientific workflow is bound to run for a considerably long time
period. The Condor software system provides scheduling mechanisms to access the
low-layer grid resources. A grid workflow enactor must provide interface to
access the services of these external scheduling and middleware systems.

\begin{table}
\begin{center}
%\hspace*{-1cm}
% use packages: array
\begin{tabular}{|l|l|l|}
\hline
\textbf{SWE/Language} & \textbf{DCI Interfacing Mechanisms} & \textbf{Currently Interfaced DCIs}\\
\hline
Taverna/Scufl2 & By Extension & None Inherently \\\hline
+/BPEL & Ad-hoc & None Inherently \\\hline
Triana/- & Provides Adapters & P2P Networks, NGS (ngs.org.uk)  \\\hline
MOTEUR/Scufl & Inherent & EGI (www.eu-egi.eu), g5k (www.grid5000.fr)\\\hline
Kepler/MoML & Using Directors & EarthGrid (http://seek.ecoinformatics.org) \\\hline
GridNexus/GXPL & Inherent & Generic Grids \\\hline
Swift/SwiftScript & Inherent & IBM/bluegene/P HPC Clusters \\\hline
Askalon/AGWL & Inherent & Austrian Grid (http://www.austriangrid.at/)\\\hline
P-GRADE/- & Inherent & NGS \\\hline
Pegasus/DAX & Using Condor & OSG (www.opensciencegrid.org) \\\hline
Galaxy/- & Extensions & None Inherently \\\hline
MOTEUR2/Gwendia~ & Inherent & EGI, g5k, HIPernet Cloud \\\hline
\end{tabular}
%\vspace*{1cm}
%\caption{A summary of SWEs DCI Interfacing Mechanisms}
\label{tbl:interfacedci}
\end{center}
\end{table}

Remotely located grid resources are accessed normally over the Internet links.
In such situations, long network latency or outright link failures are common.
Workflow managers are likely to face issues such as no response from the job
submitter or long delays. There must be some mechanism in place to handle such
scenarios. Depending upon the sophistication of the underlying infrastructure a
failover mechanism or logging information could be obtained for failures. Often
times workflow managers have their own failover mechanisms in place.

In the rest of this section, we study the approaches and techniques of SWEs to
interact with the Computing and Data Infrastructures. We also discuss briefly
the mechanisms for the mandatory security requirements.

MOTEUR provides seamless access to local and remote grid data resources by
closely coupling itself with the vBrowser~\cite{olabarriaga-deboer-etal:2006}
virtual resource browser environment. It has a built-in job management
mechanism that includes building job-specification, job submission data staging
and collection of results from the jobmanager. Additionally, MOTEUR2 is
interfaced with the HiPerNet cloud, a virtual platform manager that takes into
account, both computing and network resources
\cite{vicat-blancprimet-roca-etal:2009}.

In Taverna, a processor's invocation sequence can be customized
programmatically to enhance the processor's performance selectively. This
feature can be exploited by effectively interfacing a Taverna processor to Grid
middleware and create jobs and inputs/outputs definitions for that middleware.

In Taverna, a processor's invocation sequence can be customized
programmatically to enhance the processor's performance selectively. This
feature can be exploited by effectively interfacing a Taverna processor to Grid
middleware and create jobs and inputs/outputs definitions for that middleware.

Pegasus interfaces explicitly and transparently with the infrastructure in that
it maps the simplified workflow DAGs to the grid resources through the Condor
resource management service. Pegasus interacts with the available replica
management services in order to optimize the data access cost for the workflow
executions at runtime. It does cluster--group similar jobs together, in order
to ease the load on the scheduler and the dispatching of those jobs to the
worker nodes.

Swift/Karajan and Pegasus interface to a POSIX-style shared file system which
is one crucial aspect to their performance. %== may be shifted to data desc ==

Askalon does provide advanced resources mapping and reservations. These
features however are limited to the fact that not all underlying
infrastructures allow for a resource reservation scheme as offered by ASKALON.
GridNexus uses the OGSA-DAI (Open Grid Services Architecture Data Access and
Integration) services to access Globus grid data resources. It uses
web/grid-service clients in order to interact with the computations wrapped by
those services.

The data management services of Trident supports local storage as well as the
remote data storage services such as the Amazon S3
(\textit{http://aws.amazon.com/s3}) and SQL Server Data Services (SSDS)
(\textit{http://microsoft.com/sql/dataservices}) clouds.

Table \ref{tbl:interfacedci} presents a summary of the studied SWEs mechanisms for interfacing DCIs.