### 2014-06-02 text/scale.tex

\section{Scalability and Optimization} \label{sec:scal}
A simple definition of scalability in the context of SWE is that the
performance of the workflow engine increases linearly with the increasing
number of similar tasks. Depending upon the lag or increase in the linearity we
say a workflow enactor scales up or down.  In this section, we focus on
the scalability aspects of the workflow runtime. We do not consider the
specific scalability issues related to a distributed infrastructure such as
network latencies and protocol related overheads. However, there have been efforts to
thwart such issues at the workflow level by combining workflow enactment
techniques with those of distributed computing.  One such example is the
map/reduce framework \cite{mapreduce}.  Recently, the map/reduce framework,
that aims at performing distributed processing on large datasets, has gained a
significant popularity in workflow community. It is based on the master/worker
metaphor of distributed processing.  It employs the commonly used \textit{map}
and \textit{reduce} functions from the functional programming languages. The
map function chops a large chunk of data into smaller pieces and distributes it
to \textit{workers} which in turn can also chop it to further smaller datasets
forming a tree-like hierarchy. The worker process the data and return the
results back to their parents. The reduce step collects the results from the
workers and combines them to obtain the final result. One of the main reason
for the popularity of the map/reduce framework for workflows is the recent
attention to collections based datasets and processing. There exists
implementations of the map/reduce framework written in different languages with
enhancements and support services such as a sophisticated query service. Hadoop
(\textit{http://hadoop.apache.org/}) is one such implementation which has
already been interfaced by the Kepler SWE~\cite{wang-crawl-etal:2009}.

Large-scale applications with thousands of tasks often pose memory and disk
management difficulties leading to scalability issues.  However, the
performance of a workflow system is dependent on several factors including the
type of tasks and approach for submitting tasks for execution.  There is not
much research done towards the enactor scalability baring a couple of studies
carried out by C. Stratan et. al. \cite{stratan-iosup-etal:2008} and Deelman
et. al. \cite{callaghan-deelman-etal:2010}.  One way to compare the scalability
of different workflow environments is to experiment with a similar set of
workflows on a given system. In addition, different SWEs interface to different
DCIs. This is a time consuming process.  In the current study we do not
directly compare the workflow system scalability but we study the similarities
and differences in the techniques employed by workflow engines in order to
achieve scalability.

The software architecture of the system plays a key role in determining the
scalability for the workflow applications exhibiting a large number of tasks.
An architecture that takes into account the properties of the underlying
infrastructure could scale well in the face of growing number of workflow
tasks.

In the context of workflow runtime, scalability can be further classified in
terms of data motion scalability. It can be expressed as the rate of increase
in number of data movement quantity with respect to the increase in the elapsed
time.

While the computational scalability of a large scale experiment depends upon
the underlying computing infrastructure, at the workflow enactor level, it
becomes important to be able to prepare and dispatch as many processor
instances as possible without loss in performance. Failing to do so for a
workflow engine can become a performance bottleneck and hamper the optimal
utilization of an available computing infrastructure.

Most workflow systems exploit the parallelism at three-levels. First, at the
task-level wherein the independent branches of a workflow graph are enacted
concurrently. Second, at the data-level, where a same processor is instantiated
multiple times based upon the repeated operations on different datasets. Third,
the low-level parallelism that is encoded within the application components
independent of the workflow. At this level, the workflow enactor may
need to make a matching based upon the application type. For instance, a
workflow enactor may schedule the GPU-based processors on the resources that
indicate as possessing such capabilities.

In this section we analyze the scalability and runtime optimization measures
and techniques employed by SWEs. Illustrated below are some of the techniques
used to achieve scalability in various SWEs.

Taverna version 2.x introduced new measures in order to be able to submit a
large number of jobs simultaneously. It is able to dispatch large number of
concurrent instances of processors. An ability to implement
superpipelined~\cite{hwang:1992} execution of workflow tasks leads to a maximum
expression of parallelism~\cite{pipeline}.

Within the Taverna engine, all the data involved in the execution is managed by
a data manager through an external database, and is addressed by the engine
only by \textit{reference}. This facilitates the management of data-intensive
workflows in a scalable way, as high volume data is only transferred into the
engine space when needed by processors that are unable to resolve external data
references on their own.

Taverna, Swift-Karajan and Pegasus all provide a means to limit the maximum
number of processor instances that can run concurrently through a builtin
parameter. In Taverna, this can be configured at a workflow level while in
Pegasus, this can be configured at the Condor \cite{thain-tannenbaum-etal:2005}
scheduler level. In Swift-Karajan system this parameter is regulated
automatically as part of a ``throttling parallel foreach" construct. A similar
configurable parameter is provided by the Knime thread manager in order to
limit the size of the thread-pool employed to run a workflow.

Kepler optimizes the workflow execution through parallel and pipelined
execution. In addition to these, Kepler also provides dataflow analysis
optimization that involves data bypassing the processors in a workflow that
does not make use of that data. This is detected using a builtin Kepler
mechanism called ``static type propagation".

Another alternative is to optimize the enactment at runtime. This approach has
advantages including a better information about the actual data available. For
instance, Askalon does runtime branch prediction, parallel loop unrolling and
loop sequentialization.

The Pegasus workflow manager introduced subworkflows because of scalability
issues at the localhost for a large number of nodes in a DAG. A hierarchical
representation helps it scale at the local planning level as well as at the
execution level on distributed resources.

Static planners in Pegasus translate a workflow graph into a location specific
DAGMan input file, adding stages for data staging, intersite transfer and data
registration. These planners can reduce processing of tasks for files that
already exist, select sites for jobs, and cluster jobs based on various
criteria. Pegasus performs graph transformation with the knowledge of the whole
workflow graph. A similar strategy is grouping tasks by application-level
wrappers written in higher-level programming languages such as bash or Python.

In some systems the structure of a workflow is constructed and expanded
dynamically as opposed to static planning. A static workflow graph allows
optimizations based on the global state of the workflow, whereas a dynamic
workflow structure allows adaptations to changing environments.  Depending upon
the execution environment and access conditions to the available resources,
each approach can result in the overall optimization of a given application.

Knime has a different execution strategies for shared memory Symmetric
Multiprocessing Systems (SMP) and distributed memory systems. In the case of
SMP, it uses the message passing library or MPI to distribute the independent
parts of a workflow to available processors. This approach results in efficient
use of available processors. However, for a large scale, the global memory
becomes a bottleneck hindering the scalability.

In the case of distributed memory and independent computational nodes connected
with highspeed network or the internet, Knime employs a Master/Worker metaphor
of computing. A Knime installation with GUI acts as a master node that
coordinates the bare Knime installations on the worker nodes. These worker
nodes are then sent the portions of a workflow by the master node. The master
node keeps track of the participating nodes by their unique IP addresses. At
the end of each execution, the master node collects the results from the worker
nodes.

\begin{table}[ht!]
%TODO: Caption is not printing well, fix
%\caption{A summary of SWEs Scalability and optimization Mechanisms}
\begin{center}
% use packages: array
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\textbf{SWE} / & \multicolumn{3}{c|}{\textbf{Parallelism}} & \multicolumn{2}{c|}{\textbf{Optimization}}\\
\cline{2-6}
\textbf{Language} & \textbf{data} & \textbf{task} & \textbf{pipeline} & \textbf{static} & \textbf{runtime}\\
\hline
Taverna/Scufl2 & \checkmark & \checkmark & \checkmark & -- & \checkmark \\
\hline
Triana/- & \checkmark & \checkmark & \checkmark & -- & \checkmark \\
\hline
MOTEUR/Scufl & \checkmark & \checkmark & \checkmark &-- &\checkmark \\
\hline
Kepler/MoML & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
\hline
GridNexus/GXPL & -- & \checkmark & -- &-- &\checkmark \\
\hline
Swift/SwiftScript & \checkmark & \checkmark & \checkmark & \checkmark &\checkmark \\
\hline
Askalon/AGWL &--& \checkmark & -- & \checkmark & \checkmark \\
\hline
P-GRADE/- &\checkmark &\checkmark &\checkmark &--&--\\
\hline
Pegasus/DAX &\checkmark &\checkmark &\checkmark &\checkmark &\checkmark \\
\hline
Galaxy/- &-- &-- &\checkmark &-- &-- \\
\hline
MOTEUR2/Gwendia~ & \checkmark &\checkmark & \checkmark &-- &\checkmark \\
\hline
\end{tabular}
\label{tbl:scalability}
\end{center}
\end{table}

Shown in Table \ref{tbl:scalability} is a summary of the Scalability and
optimization mechanisms of the studied SWEs. The scalability features
considered are through three levels of parallelism in the form of data,
task and pipeline. The optimization features considered are the presence of
mechanisms by which static and runtime optimizations are achieved.