### 2014-06-02 paper.tex
hpc vs htc

% v2-acmsmall-sample.tex, dated March 6 2012
% This is a sample file for ACM small trim journals
%
% Compilation using 'acmsmall.cls' - version 1.3 (March 2012), Aptara Inc.
% (c) 2010 Association for Computing Machinery (ACM)
%
% Questions/Suggestions/Feedback should be addressed to => "acmtexsupport@aptaracorp.com".
% Users can also go through the FAQs available on the journal's submission webpage.
%
% Steps to compile: latex, bibtex, latex latex
%
% For tracking purposes => this is v1.3 - March 2012

\documentclass[prodmode,acmtecs]{acmsmall} % Aptara syntax

% Package to generate and customize Algorithm as per ACM style
\usepackage[ruled]{algorithm2e}
\usepackage{cite}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{mindmap,trees}
\renewcommand{\algorithmcfname}{ALGORITHM}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}
\setlength\parindent{0pt}
\usepackage{setspace}
%\doublespacing
\linespread{2}

% Metadata Information
\acmVolume{9}
\acmNumber{4}
\acmArticle{39}
\acmYear{2010}
\acmMonth{3}

\newcommand{\jm}[1]{
    \noindent{\textcolor{red}{\bf\em #1 -- Johan}}
}
\newcommand{\km}[1]{
    \noindent{\textcolor{blue}{\bf\em #1 -- Ketan}}
}

% Document starts
\begin{document}

% Page heads
\markboth{K. Maheshwari et al.}{A Survey of Scientific Workflow Management Systems}

% Title portion
\title{State of the Art in Scientific Workflow Systems}
\author{KETAN MAHESHWARI
\affil{University of Nice, Sophia Antipolis}
JOHAN MONTAGNAT
\affil{CNRS}
}
% NOTE! Affiliations placed here should be for the institution where the
%       BULK of the research was done. If the author has gone to a new
%       institution, before publication, the (above) affiliation should NOT be changed.
%       The authors 'current' address may be given in the "Author's addresses:" block (below).
%       So for example, Mr. Abdelzaher, the bulk of the research was done at UIUC, and he is
%       currently affiliated with NASA.

\begin{abstract}

\km{TODO: Expand in the end}

In this work we conduct a literature survey of the key features of the
contemporary scientific workflow environments. We define the criteria for the
survey based chiefly on user needs arising from diverse applications and
infrastructure. An overview of studied scientific workflow environments is
presented followed by the identification and description of the features offered by
each of the environment. A tabulated summary for each feature class gives a
high-level view of the offered features. Finally, a taxonomy is proposed
placing these features into respective places in a hierarchy.

\jm{Test comment}

\end{abstract}
%\category{C.2.2}{Computer-Communication Networks}{Network Protocols}
%\terms{Design, Algorithms, Performance}
%\keywords{Wireless sensor networks, media access control, multi-channel, radio interference, time synchronization}

\acmformat{Ketan Maheshwari, Johan Montagnat
, 2014. A survey of scientific workflow management systems.}
% At a minimum you need to supply the author names, year and a title.
% IMPORTANT:
% Full first names whenever they are known, surname last, followed by a period.
% In the case of two authors, 'and' is placed between them.
% In the case of three or more authors, the serial comma is used, that is, all author names
% except the last one but including the penultimate author's name are followed by a comma,
% and then 'and' is placed before the final author's name.
% If only first and middle initials are known, then each initial
% is followed by a period and they are separated by a space.
% The remaining information (journal title, volume, article number, date, etc.) is 'auto-generated'.

\begin{bottomstuff}
This work is supported by the ANR-CNRS, under
grant contract CNRS-0435060.

Author's addresses: Ketan Maheshwari, Argonne National Laboratory; Johan Montagnat, I3S, Sophia Antipolis.
\end{bottomstuff}

\maketitle

%Workflows considered:
% MOTEUR2
% Taverna
% Swift
% Pegasus
% Triana % not sure
% Galaxy
% Makeflow
% Kepler/Ptolemy
% Spark
% Linda
% DaGue
% Wings
% PGRADE Portal

\input{text/intro}
\input{text/concepts}
\input{text/related}
\input{text/objectives}

\section{Overview of the Workflow Environments} \label{sec:overview}
In this section we present an overview of the SWEs studied. The paragraph
headers are organized by the name of the SWE followed by the name of internal
representation language use and support separated by a forward slash (/).  The
cases where an intermediate language is not specifically defined or conversely
where there are multiple systems supporting a generic representation are
denoted by an asterisk (*) or a hyphen (-) at the respective places.

\input{text/taverna}
\input{text/moteur}
\input{text/makeflow}
\input{text/triana}
\input{text/kepler}
\input{text/peg}
\input{text/wings}
\input{text/swift}
\input{text/askalon}
\input{text/pgrade}
\input{text/galaxy}
\input{text/others}
% Put this in analysis section. 
%The \textit{gscript}~\cite{maheshwari-montagnat:2010} workflow language
%is similar in spirit to the SwiftScript. Similar to SwiftScript, gscript adopts futures to express
%asynchronous execution in a script-based language. A Swift script 
%contains statically typed datasets that can be declared
%in the script. Unlike Swift script, gscript does provide user-defined
%strategies in order to effect data combinations in the workflow. The underlying datasets organization in Swift is
%represented using XML-representation. This means that effectively, a
%workflow is a DAG. A workflow is
%simpler and compact in gscript as its data description is embedded within a
%workflow representation. Conditional and Loop execution is supported
%using the if and foreach statements in both languages.
%
\input{text/scale}
\input{text/data}
\input{text/interface}
\input{text/lang}
\input{text/usability}
%
\input{text/concl}
% Bibliography
\bibliographystyle{ACM-Reference-Format-Journals}
\bibliography{sanebib}
%\nocite{*}
\end{document}